Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy

### 2023-12-09 Text/Articles/MIGARS_2024/R1-The Entropy as a Proxy for Fully Developed Speckle.Rmd
adding improvements for code and paper

---
title: Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy
affiliation:
  ## Author mode :  use one only of the following --- +

  ## one column per author - using only the institution field
  author-columnar: true # uncomment this line to use

@@

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{mathtools}
   - \usepackage[utf8]{inputenc}
   - \usepackage{hyperref}
   - \hypersetup{draft}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{mathabx}

bibliography:  ../../Common/references.bib
csl: "ieee.csl"
output: 
  rticles::ieee_article:
    number_sections: TRUE

#citation_sorting: none   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# library(ggplot2)
# library(ggthemes)
# library(reshape2)
# library(plotly)
# library(knitr)
# # library(rgl)
# library(gtools)
# library(stats4)
# library(rmutil)
# library(gsl)

library(ggplot2)

library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(invgamma)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(ggsci)
library(wesanderson)
library(ggpubr)
library(patchwork)
library(dplyr)
options(kableExtra.latex.load_packages = FALSE)
library(devtools)
devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")# Gridtop , right , bottom , or left#, panel.grid = element_blank()
)

# theme_set(theme_pander() +
#             theme(text=element_text(family="serif"),
#                   legend.position = "top")
#           )

@@

Synthetic aperture radar (SAR) has become a fundamental technology for environmental monitoring and disaster management because of its ability to provide daytime and nighttime imagery in all weather conditions&nbsp;[@Mu2019]. 
However, the utility of SAR data depends on a thorough understanding of their statistical properties.
Speckle is part of SAR data because of the imaging process' coherent nature.
Its non-additivity and non-Gaussianity require robust statistical models that can accurately characterize the data.

@@

When deciding which model is the best, practitioners face a problem.
On the one hand, if they opt for the Gamma law when the data come from the $\mathcal{G}^0$ distribution, they lose all the information about the number of scatterers, which is revealed by one of the parameters of the latter model&nbsp;[@Yue2021].
On the other hand, if they apply the $\mathcal{G}^0$ distribution under fully developed speckle,
maximum likelihood estimation is tricky: 
bias increases making estimation unreliable&nbsp;[@VasconcellosFrerySilva:CompStat], and
the likelihood is flat, so numerical optimization may not converge&nbsp;[@FreryCribariSouza:JASP:04].
The two-stage technique we propose tackles this problem by using the entropy as a proxy to decide which is the best model.

@@

Estimating the entropy faces practical challenges, particularly when the model is unknown; non-parametric methods are utilized in such cases.
Among non-parametric approaches,&nbsp;@Subhash2021 discussed the use of spacing methods.
This non-parametric strategy offers flexibility to address a wide range of models without imposing specific parametric constraints. We extend the exploration of non-parametric entropy estimators by incorporating enhanced bootstrap methodologies.

@@

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully developed speckle and is a limiting case of the second, which is appealing due to its versatility in accurately representing regions with various roughness characteristics&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim G_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:
\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align}
where, in&nbsp;\eqref{E:gamma1} $\mu > 0$ is the mean; in&nbsp;\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, $\Gamma(\cdot)$ is the gamma function, and $\mathbbm 1_{A}(z)$ is the indicator function of the set $A$.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$&nbsp;[@Nascimento2010].
In this way, we compute the expected value $\mu$ using the expression in&nbsp;\eqref{E:rmom}, and we reparametrize&nbsp;\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density function that characterize the $G_I^0(\mu, \alpha, L)$ law is
\begin{multline}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)}\\ \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{multline}

@@

Using&nbsp;\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$in&nbsp;\eqref{E:gamma1} and $\mathcal{G}_I^0$in&nbsp;\eqref{E:gi02} based on @Cassetti2022 and @Ferreira2020:
\begin{align*}
H_{\Gamma_{\text{SAR}}}(L, \mu) &= L - \ln L + \ln\Gamma(L) + (1-L)\psi^{(0)}(L) + \ln \mu, \\
H_{\mathcal{G}_I^0}(\mu, \alpha, L) &= L - \ln L + \ln\Gamma(L) + (1-L)\psi^{(0)}(L) + \ln \mu \\
&\quad - \ln\Gamma(L-\alpha) + (L-\alpha)\psi^{(0)}(L-\alpha)- L\\
&\quad - (1-\alpha)\psi^{(0)}(-\alpha) + \ln (-1-\alpha) + \ln\Gamma(-\alpha)
\end{align*}
where $\psi^{(0)}(\cdot)$ is the digamma function. In Fig.&nbsp;\ref{fig:PlotGammaSAR} we see how the Entropy of the Gamma SAR distribution changes with $\mu$ for various values of $L$.

```{r FunctionsDefinitions, echo=FALSE}
entropy_gamma_sar <- function(L, mu) {

    return(L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu))

}

entropy_gI0 <- function(mu, alpha, L) {

  term1 <- L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu)   
  term2 <- -L - log(gamma(L-alpha)) + (L-alpha)*(digamma(L - alpha))- (1-alpha)*digamma(- alpha)+log(-1 - alpha)+log(gamma(-alpha))

  entropy <- term1 + term2 
  return(entropy)
}
```

@@

One of the earliest non-parametric estimators relying on spacings was introduced by @vasicek1976test. 
Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* @correa1995new: $\widehat{H}_{\text{C}}$.

* @Ebrahimi1994: $\widehat{H}_{\mathrm{E}}(\bm{Z})$.

* @IbrahimAlOmari2014: $\widehat{H}_{\mathrm{AO}}(\bm{Z})$.

These estimators, along with others, are described and studied in @Cassetti2022.

<!-- * Correa [@correa1995new]: -->
<!-- \begin{equation*} -->
<!-- \widehat{H}_{\text{C}}(\bm{Z}) = -\frac{1}{n} \sum_{i=1}^n \ln \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\widebar{Z}_{(i)}\right)}{n \sum_{j=i-m}^{i+m}\left(Z_{(j)}-\widebar{Z}_{(i)}\right)^2}, -->
<!-- \end{equation*} -->
<!-- where $\widebar{Z}_{(i)} = (2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}$. -->

<!-- * Ebrahimi [@Ebrahimi1994]: -->
<!-- \begin{equation*} -->
<!-- \widehat{H}_{\mathrm{E}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \ln \left[\frac{n}{\tau_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right], -->
<!-- \end{equation*} -->
<!-- where -->
<!-- $$ -->
<!-- \tau_i= \begin{cases}1+(i-1) / m & \text { if } 1 \leq i \leq m, \\ 2 & \text { if } m+1 \leq i \leq n-m, \\ 1+(n-i) / m & \text { if } n-m+1 \leq i \leq n.\end{cases} -->
<!-- $$ -->

<!-- * Al-Omari [@IbrahimAlOmari2014]: -->
<!-- \begin{equation*} -->
<!-- \widehat{H}_{\mathrm{AO}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \ln \left[\frac{n}{\omega_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right], -->
<!-- \end{equation*} -->
<!-- where -->
<!-- $$ -->
<!-- \omega_i= \begin{cases}3 / 2 & \text { if } 1 \leq i \leq m, \\ 2 & \text { if } m+1 \leq i \leq n-m, \\ 3 / 2 & \text { if } n-m+1 \leq i \leq n.\end{cases} -->
<!-- $$ -->

@@

# Hypothesis testing based on non-parametric entropy {#sec:test}
General asymptotic results for functions of spacings are detailed&nbsp;@Khashimov1990, while&nbsp;@Bert1992 developed a correction for the case of Shannon entropy.
Following the work of these authors, the next result applies: 
\begin{lemma}
Suppose that $f(\cdot)$ is a bounded density bounded away from zero and satisfies a Lipschitz condition on its support.
Then, if $m,n\rightarrow \infty$ and $m=o(n^{1/2})$, holds that:
\begin{equation*}
\sqrt{n}\,\Big(\label{Eq:bias_t}
\widetilde{H}_{i}+\int_{-\infty}^\infty f(z)\ln f(z) \mathrm{d}z\Big)
\xrightarrow[]{\mathcal{D}}
\mathcal{N}(0,\operatorname{Var}(\ln f(Z))).
\end{equation*}
\end{lemma} 
Consider, starting from the previous lemma, the test of the null hypothesis $\mathcal{H}_0: T_{\mathcal{D}}=D_0$ as opposed to one of the other three:
\begin{align*}
\mathcal{H}_1 &: T_{\mathcal{D}}\neq D_0,\\ 
\mathcal{H}_1 &: T_{\mathcal{D}}> D_0,\text{ or}\\
\mathcal{H}_1 &: T_{\mathcal{D}}< D_0. 
\end{align*}
For this purpose, we can use the test statistics:
\begin{equation*}
Z_{m,n} = \frac{\sqrt{n}\big(\label{Eq:bias4}
\widetilde{H}-D_0\big)}{\sqrt{\operatorname{Var}\big(\ln f(Z)\big)}},
\end{equation*}
so the null hypothesis should be rejected if (i) $Z_{n,m} > z_{\alpha/2}$ or $Z_{n,m} < - z_{\alpha/2}$ for $\Phi_{\mathcal N}(z_{\alpha/2})=1-\alpha/2$ and $\Phi_{\mathcal N}$ being the standard normal cumulative distribution function, (ii) $Z_{n,m} > z_{\alpha/2}$ or (iii) $Z_{n,m} < - z_{\alpha/2}$.

@@

Simulations are conducted using $\mathcal{G}_I^0$ distribution, with $200$ simulated samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\), with parameters \(\mu \in \left\{1, 10\right\}\), \(\alpha=-20\), and \(L=8\). In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $\\mathcal{G}_I^0$, $L=8$, $\\alpha=-20$.", fig.width=6, fig.height=5}
# Define sample sizes for analysis
set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 200

# Number of bootstrap replications
B <- 100
mu_values <- c(1, 10)
alpha <- -20
L <- 8
# \widehat{H}_{\text{C}}, \widehat{H}_{\text{E}} \widehat{H}_{\text{AO}}, \widehat{H}_{\text{V}}
# \widetilde{H}_{C}, \widetilde{H}_{E}, \widetilde{H}_{AO}
estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

combined_plot <- generate_plot_gi0(sample_sizes, R, B, mu_values,alpha, L, estimators,  ncol = 1, nrow = 2)

# Print the combined plot with sanitize=FALSE
print(combined_plot)

@@

```{r Table_gi01, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#
set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)
R <- 200
B <- 100
mu_values <- c(1,  10)
L <- 8
alpha<- -20

# Define a list of estimators
estimators <- list(
  #"Van Es Bootstrap" = bootstrap_van_es_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
  #"Vasicek Bootstrap" = bootstrap_vasicek_estimator,
  "Al Omari  Bootstrap" = bootstrap_al_omari_1_estimator
  #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
)

# 
combined_results <- NULL

for (mu_val in mu_values) {
  # 
  results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val,alpha,  L, estimators)

  # 
  results <- cbind(mu = mu_val, results)

  # 
  combined_results <- bind_rows(combined_results, results)
}

# 
reshaped_results <- combined_results %>%
  pivot_wider(names_from = Estimator, values_from = c(Bias, MSE))
#\widetilde{H}_{i}

colnames(reshaped_results) <- c("$\\bm{\\mu}$", "$\\bm{n}$",  "$\\widetilde{H}_{C}$", "$\\widetilde{H}_{E}$",  "$\\widetilde{H}_{AO}$",  "$\\widetilde{H}_{C}$", "$\\widetilde{H}_{E}$",  "$\\widetilde{H}_{AO}$")

print(
  kbl(
    reshaped_results,
    caption = "Bias and MSE of bootstrap estimators for $\\mathcal{G}_I^0$, $L=8$, $\\alpha=-20$.",
    format = "latex",
    booktabs = TRUE,
    align = "c",
    escape = FALSE, 
    digits = 4,  
    label = "table2"  # 
  ) %>%
    add_header_above(c(" ", " ", "Bias" = 3, "MSE" = 3)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")# %>%
  %>%
    kable_styling(latex_options = c( "striped")) 
   # column_spec(1, bold = TRUE) 
)

@@

@@

# Tamaños de muestra
sample_sizes <- c(9, 25, 49)
R <- 100
mu <- 10
L <- 5
alpha <- -400

# Entropía verdadera de Gamma SAR
true_entropy <- entropy_gamma_sar(L, mu)
#
# Función para generar muestras
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)#
  }
  return(samples)
}

>>>>

---
title: Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy
affiliation:

  author-columnar: true # uncomment this line to use

@@

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{mathtools}
   - \usepackage[utf8]{inputenc}
   - \usepackage{hyperref}
   - \hypersetup{draft}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{mathabx}
  # - \usepackage{pgf}
   - "\\setcitestyle{square,comma,numbers,sort&compress}" # citation style to numerical, work with \cite{} or [@]

output:
  rticles::ieee_article:
    number_sections: TRUE
    #citation_package: biblatex
    citation_package: natbib

bibliography: ../../Common/references.bib

biblio-style: ieeetr

#link-citations: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
#library(RColorBrewer)
library(ggsci)
#library(wesanderson)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

@@

Synthetic aperture radar (SAR) has become a fundamental technology for environmental monitoring and disaster management because of its ability to provide daytime and nighttime imagery in all weather conditions&nbsp;\cite{Mu2019}. 
However, the utility of SAR data depends on a thorough understanding of their statistical properties.
Speckle is part of SAR data because of the imaging process' coherent nature.
Its non-additivity and non-Gaussianity require robust statistical models that can accurately characterize the data.

@@

When deciding which model is the best, practitioners face a problem.
On the one hand, if they opt for the Gamma law when the data come from the $\mathcal{G}^0$ distribution, they lose all the information about the number of scatterers, which is revealed by one of the parameters of the latter model&nbsp;\cite{Yue2021}.
On the other hand, if they apply the $\mathcal{G}^0$ distribution under fully developed speckle,
maximum likelihood estimation is tricky: 
bias increases making estimation unreliable&nbsp;\cite{VasconcellosFrerySilva:CompStat}, and
the likelihood is flat, so numerical optimization may not converge&nbsp;\cite{FreryCribariSouza:JASP:04}.
The two-stage technique we propose tackles this problem by using the entropy as a proxy to decide which is the best model.

@@

Estimating the entropy faces practical challenges, particularly when the model is unknown; non-parametric methods are utilized in such cases.
Among non-parametric approaches,&nbsp;\cite{Subhash2021} discussed the use of spacing methods.
This non-parametric strategy offers flexibility to address a wide range of models without imposing specific parametric constraints. We extend the exploration of non-parametric entropy estimators by incorporating enhanced bootstrap methodologies.

@@

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;\cite{Frery1997}. 
The first is suitable for fully developed speckle and is a limiting case of the second, which is appealing due to its versatility in accurately representing regions with various roughness characteristics&nbsp;\cite{Cassetti2022}.
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:
\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align}
where, in&nbsp;\eqref{E:gamma1} $\mu > 0$ is the mean; in&nbsp;\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, $\Gamma(\cdot)$ is the gamma function, and $\mathbbm 1_{A}(z)$ is the indicator function of the set $A$.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	\mathbbm E_{\mathcal{G}_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$&nbsp;\cite{Nascimento2010}.
In this way, we compute the expected value $\mu$ using the expression in&nbsp;\eqref{E:rmom}, and we reparametrize&nbsp;\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density function that characterize the $G_I^0(\mu, \alpha, L)$ law is
\begin{multline}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)}\\ \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{multline}

@@

Using&nbsp;\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$in&nbsp;\eqref{E:gamma1} and $G_I^0$in&nbsp;\eqref{E:gi02} based on&nbsp;\cite{Cassetti2022} and&nbsp;\cite{Ferreira2020}:
\begin{multline}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{multline}
\begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu \\
-\ln\Gamma(L-\alpha)+ (L-\alpha) \psi^{(0)}(L-\alpha)\\
-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline}
where $\psi^{(0)}(\cdot)$ is the digamma function.

In Fig.&nbsp;\ref{fig:PlotGammaSAR} we see how the Entropy of the Gamma SAR distribution changes with $\mu$ for various values of $L$.

@@

One of the earliest non-parametric estimators relying on spacings was introduced by&nbsp;\cite{vasicek1976test}. 
Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* \cite{correa1995new}: $\widehat{H}_{\text{C}}$.

* \cite{Ebrahimi1994}: $\widehat{H}_{\mathrm{E}}$.

* \cite{IbrahimAlOmari2014}: $\widehat{H}_{\mathrm{AO}}$.

These estimators, along with others, are described and studied in&nbsp;\cite{Cassetti2022}.

@@

# Hypothesis testing based on non-parametric entropy {#sec:test}
General asymptotic results for functions of spacings are detailed&nbsp;\cite{Khashimov1990}, while&nbsp;\cite{Bert1992} developed a correction for the case of Shannon entropy.
Following the work of these authors, the next result applies: 
\begin{lemma}
Suppose that $f(\cdot)$ is a bounded density bounded away from zero and satisfies a Lipschitz condition on its support.
Then, if $m,n\rightarrow \infty$ and $m=o(n^{1/2})$, holds that:
\begin{equation*}
\sqrt{n}\,\Big(\label{Eq:bias_t}
\widetilde{H}_{i}+\int_{-\infty}^\infty f(z)\ln f(z) \mathrm{d}z\Big)
\xrightarrow[]{\mathcal{D}}
\mathcal{N}(0,\operatorname{Var}(\ln f(Z))).
\end{equation*}
\end{lemma} 
Consider, starting from the previous lemma, the test of the null hypothesis $\mathcal{H}_0: T_{\mathcal{D}}=D_0$ as opposed to one of the other three:
\begin{align*}
\mathcal{H}_1 &: T_{\mathcal{D}}\neq D_0,\\ 
\mathcal{H}_1 &: T_{\mathcal{D}}> D_0,\text{ or}\\
\mathcal{H}_1 &: T_{\mathcal{D}}< D_0. 
\end{align*}
For this purpose, we can use the test statistics:
\begin{equation*}
Z_{m,n} = \frac{\sqrt{n}\big(\label{Eq:bias4}
\widetilde{H}-D_0\big)}{\sqrt{\operatorname{Var}\big(\ln f(Z)\big)}},
\end{equation*}
so the null hypothesis should be rejected if (i) $Z_{n,m} > z_{\alpha/2}$ or $Z_{n,m} < - z_{\alpha/2}$ for $\Phi_{\mathcal N}(z_{\alpha/2})=1-\alpha/2$ and $\Phi_{\mathcal N}$ being the standard normal cumulative distribution function, (ii) $Z_{n,m} > z_{\alpha/2}$ or (iii) $Z_{n,m} < - z_{\alpha/2}$.

@@

Simulations are conducted using $G_I^0$ distribution, with $200$ simulated samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\), with parameters \(\mu \in \left\{1, 10\right\}\), \(\alpha=-20\), and \(L=8\). In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

```{r Simulated_data_gi0, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 10

# Number of bootstrap replications
B <- 10
mu_values <- c(1, 10)
alpha <- -20
L <- 8

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/results_gi0.Rdata")

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=8$, $\\alpha=-20$.", fig.width=6, fig.height=5}

load("./Data/results_gi0.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{H}_{C}$"),#parse(text=TeX("$H_{2}$")),
                             "Correa Bootstrap" = expression("$\\widetilde{H}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{H}_{E}$"),
                            "Al Omari" = expression("$\\widehat{H}_{AO}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{H}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{H}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 2)

print(combined_plot_gi0)

@@

```{r Table_gi0, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_gi0.Rdata")

estimators_to_table <- c("Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap")

filtered_results <- purrr::map_dfr(results_gi0, ~ .x %>% filter(Estimator %in% estimators_to_table), .id = "mu")

reshaped_results <- filtered_results %>%
  pivot_wider(names_from = Estimator, values_from = c(Bias, MSE))

colnames(reshaped_results) <- c("$\\bm{\\mu}$", "$\\bm{n}$", "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$", "$\\widetilde{H}_{\\text{AO}}$", "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$", "$\\widetilde{H}_{\\text{AO}}$")

#
print(
  kbl(
    reshaped_results,
    caption = "Bias and MSE of bootstrap estimators for $G_I^0$, $L=8$, $\\alpha=-20$.",
    format = "latex",
    booktabs = TRUE,
    align = "c",
    escape = FALSE, 
    digits = 4,  
    label = "table2"  
  ) %>%
    add_header_above(c(" ", " ", "Bias" = 3, "MSE" = 3)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")
)
```

@@

sample_sizes <- c( 9, 25, 49, 81, 121)
R <- 100
mu <- 5
L <- 2
alpha <- -1000

true_entropy <- entropy_gamma_sar(L, mu)

generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)#
  }
  return(samples)
}
